---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

# devtools::build_readme()

```

# TDAtools

<!-- badges: start -->
<!-- badges: end -->

TDAtools provides many functions to help data scientists using TDA.

## Installation

You can install the development version of TDAtools from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
devtools::install_github("vituri/TDAtools")
```

## Notation

Let's fix some notation:

- `X` is a data.frame with numeric and, possibly, factor columns. When needed, we select only the numeric columns (like when calculating the distance matrix of `X`) and denote it by $X$. The rows of `X` are called observations, and the columns are called variables;

- `f_X` is a vector with one number for each observation of `X`. That is: `f_X` is the result of applying a *filter function* $f: X \to \mathbb{R}$ to `X`. We provide many filter functions;

- `distance_matrix` is a matrix of distances calculated with `X` OR a function that when applied to `X` results in such a matrix

## Graph-reducing algorithms

We provide the Mapper algorithm and the Ball-Mapper algorithm. Both return a list consisting of:

- a graph with weighted vertices;

- a list showing which points of `X` are in each vertex.

See the vignettes for more examples. (to-do)

## Examples

Let `X` be a sample of random points in a noisy circle:

```{r example}
library(TDAtools)

X = data.noisy_circle()

```

Let the filter function be the projetion on the x-axis and let color `X` using it:

```{r}
f_X = X$x

color_band = ggplot2::cut_interval(f_X, n = 50)
col_vector = viridis::viridis(nlevels(color_band))[as.integer(color_band)]
plot(X, col = col_vector)
```

Now we embed `X` in $R^4$:

```{r}
X$z = 0
X$w = 0
```

and calculate its Mapper Graph:

```{r}
mp =
  mapper(
    X = X
    ,f_X = f_X
  )
```

The result is a list with many objects. For example, the pullback of each interval in the covering:

```{r}
mp$pullback |> str()
```

Let's plot the mapper graph:

```{r}
mp$graph %>% plot()
```

Or, for a better visualization, we use the `networkVis` package and color the nodes using the mean value of the `x` variable of `X`:

```{r}
mp %>% plot_mapper()
```

Or color it by `y`:

```{r}
mp %>% plot_mapper(data_column = 'y')
```

## To do

- Vignettes using mapper with factor variables
- Ball Mapper documentation
- Persistent homology examples
- Data analysis using some vertices of the mapper + machine learning
